Moments and Minkowski Functionals constrain cosmology: applying a weak lensing emulator to CFHTLenS data

### 2015-03-03 Paper_minkowski/paper.tex
latest edits

\begin{abstract}
Weak gravitational lensing is a potentially powerful probe for cosmology, and in particular for the dark energy (DE) equation of state $w$; since typical weak lensing fields are non--Gaussian in nature, quadratic statistics, such as angular power spectra, can miss significant cosmological information contained in survey data. In this work we examine constraints on the parameter triplet $(\Omega_m,w,\sigma_8)$ derived from weak lensing Minkowski functionals and moments, using publicly available data from the 154\,deg$^2$ CFHTLenS survey. We utilize a new suite of ray--tracing N-body simulations spanning 91 points in the $(\Omega_m,w,\sigma_8)$ parameter space, replicating the galaxy sky positions, redshift and shape noise in the CFHTLenS catalogs. We then build an emulator that interpolates the simulated descriptor spaces, and use it to compute the parameter likelihood, from which we derive the data constraints. We find that the dimensionality reduction technique Principal Component Analysis helps in stabilizing the constraints with respect to the number of bins used to construct the convergence statistics. Using our full set of statistical descriptors, we constrain $\Sigma_8=\sigma_8(\Omega_m/0.27)^{0.55}$ to a value of $0.75\pm0.04$ ($1\sigma$), well in agreement with the values already quoted in the literature. We find that, while the Minkowski functionals constraints on the $(\Omega_m,\sigma_8)$ doublet suffer from a bias, the moments provide a tight unbiased bound on the parameters, with a substantial contribution coming from quartic moments of gradients.

@@

Weak gravitational lensing is emerging as a promising technique to constrain cosmology. Techniques have been developed to construct cosmic shear fields with shape measurements in large galaxy catalogues; although the shear two--point function (2PCF) serves as a main cosmological probe (see \citep{CFHTKilbinger} for example), alternative statistics have been shown to increase the amount of cosmological information one can extract from weak lensing fields. Among these, higher moments (\citep{moments1,moments2,moments3,moments4,moments5,moments6}), three--point functions (\citep{3pcf1,3pcf2}), bispectra (\citep{bispectrum1,bispectrum2,bispectrum3,bispectrum4}), peak counts (\citep{peaks1,peaks2,peaks3,peaks4,peaks5,peaks6}) and Minkowski Functionals (\citep{MinkJan,Petri2013}) have shown to improve cosmological constraints in weak lensing analyses.  
In this work we use the publicly available CFHTLenS data, consisting of a catalog of $\approx$4.2 million galaxies, combined with a suite of ray-tracing simulations in 91 different cosmological models to constrain the cosmological parameters $\Omega_m$,$\sigma_8$ and the DE equation of state $w$. The paper is organized as follows; we first give an overview of the CFHTLenS catalogs, summarizing the data reduction techniques adopted. Next, we give a description of our simulation pipeline, including the procedure used to sample the parameter space and outline the ray--tracing algorithm. We call the weak lensing statistical observables - the Power Spectrum, Minkowski Functionals and Moments - "descriptors" throughout the paper. We discuss the calculation of the statistical descriptors, the dimensional reduction method, and the statistical inference framework used. We then outline our results, and the cosmological parameter constraints obtained. To conclude, we discuss our findings and comment on possible future extensions of this analysis.

@@

\subsection{CFHTLenS data reduction}
\label{cfhtdatareduction}
%
The CFHTLenS survey covers four sky patches of 64, 23, 44 and 23 deg$^2$ area, for a total of 154 deg$^2$; the publicly released data consist of a galaxy catalog created using SExtractor \citep{SExtractor}, a photometric redshift estimated with a Bayesian photometric redshift code \citep{PhotoCode} and galaxy shape measurements made using \textit{lensfit} \citep{cfht1,cfht2}. The cosmological parameter inferences have been obtained in \citep{CFHTKilbinger} using the 2PCF; additionally a number of authors investigated the CFHTLenS constraining power using statistics that go beyond the usual quadratic ones: \citep{CFHTFu} investigated the effect of using three--point correlations as an additional probe for cosmology, while \citep{CFHTMasato} investigated CFHT cosmological constraints and systematic errors using the Minkowski Functionals. The cosmological constraints obtained in this work are based on an irregular--grid sampling of the parameter space, which is an improvement on the Fisher matrix formalism adopted by \citep{CFHTMasato}.

We apply the following cuts to the galaxy catalog: mask$<1$ (see Table B2 in \citep{SExtractor}), redshift $0.2 < z < 1.3$ (see \citep{cfht1}), fitclass = 0 (which requires the object to be a galaxy) and weight $\mathrm{w}>0$ (with larger $\mathrm{w}$ indicating smaller shear measurement uncertainty). Applying these cuts leaves us 4.2$\times10^6$ galaxies, 124.7 deg$^2$ sky coverage, and average galaxy density $n_{gal} \approx 9.3\,\mathrm{arcmin}^{-2}$. The catalog is further reduced by $\sim25\%$ when one rejects fields with non--negligible star--galaxy correlations. These spurious correlations are likely due to imperfect PSF removal, and do not contain cosmological signal. These cuts are consistent with the ones adopted by the CFHTLenS collaboration (see \citep{CFHTFu}). The CFHTLenS galaxy catalog provides us with the sky position $\pmb{\theta}$, the redshift $z(\pmb{\theta})$ and ellipticity $\mathbf{e}(\pmb{\theta})$ of each galaxy, as well as the individual weight factors $w(\pmb{\theta})$ and additive and multiplicative ellipticity corrections $c(\pmb{\theta}), m(\pmb{\theta})$. Because the CFHTLenS fields are irregularly shaped, we first divide them into 13 squares (subfields) to match the shape and $\approx12$ deg$^2$ size of our simulated maps; these square subfield maps are pixelized according to a Gaussian gridding procedure

@@

\subsection{Simulation design}
In this section we give a description of the method we used to sample the parameter space in our simulation effort. We wish to investigate the non--linear dependence of cosmological probes (in this work Minkowski Functionals and Moments of the $\kappa$ field) on the parameter triplet $\mathbf{p}=(\Omega_m,w,\sigma_8)$, while keeping the other relevant parameters $(h,\Omega_b,n_s)$ fixed to the values (0.7, 0.046, 0.96). We sampled the $D$--dimensional ($D=3$ in this case) parameter space using an irregularly spaced grid, designed with a method similar to \citep{coyote2}; the irregular grid design is more efficient, given the available computing resources: a regular parameter grid with the same average spacing between models would require a prohibitively large number of samples. We limit the parameter sampling in a box of corners $\Omega_m\in[0.07,1],\,w\in[-3.0,0],\,\sigma_8\in[0.1,1.5]$ and we map this sampling box $\Pi$ into an hypercube of unit side; we want to contruct an irregularly spaced grid consisting of $N$ points $\mathbf{x}_i\in[0,1]^D$. Let a \textit{design} $\mathcal{D}$ be the set of this irregularly spaced $N$ points: we wish to find an optimal design, in which the points are spread as uniformly as possible inside the box. Following \citep{coyote2}, we choose our optimal design as the minimum of the cost function

@@

We found that for $N=91$ grid points, order of $10^5$ iterations are sufficient to reach an accuracy of $\epsilon\sim10^{-4}$; once the optimal design $\mathcal{D}_m$ has been found, we can invert the mapping $\Pi\rightarrow[0,1]^3$ to find our simulation parameter sampling $\mathbf{p}_s$, which we show in Table \ref{designtable} and Figure \ref{designfig}.

@@

\caption{$(\Omega_m,w)$ and $(\Omega_m,\sigma_8)$ projections of our the simulation design; the blue points correspond to the CFHTemu1 simulation set, which consists of one $N$--body simulation per point, while the red point corresponds to the CFHTcov simulation set, which is based on 50 independent $N$--body simulations}

@@

For each parameter point on the grid $\mathbf{p}_s$ we run one $N$--body simulation and perform ray tracing through it, as described in \S~\ref{raysim}, to simulate CFHTLenS shear catalogs; this set of simulations will be called CFHTemu1 throughout the rest of this work. Additionally, we run 50 independent $N$--body simulations with a \textit{fiducial} parameter choice $\mathbf{p}_0=(0.26,-1.0,0.8)$ for the purpose of measuring accurately the covariance matrices which will be used for the parameter inferences in \S-\ref{cosmostats}; this set of simulations will be called CFHTcov throughout the rest of this work.

@@

The goal of this section is to give an outline of our simulation pipeline; the fluctuations in the matter density field between a source at redshift $z$ and an observer located on Earth will cause small deflections to the trajectories of light rays traveling from the source to the observer. The fluctuations in the matter field are described by the gravitational potential $\Phi(\mathbf{x},z)=\Phi(\mathbf{x}_\perp,\chi(z))$, where we can replace the physical coordinates $\mathrm{x}$ with the comoving distance from the observer $\chi$ and two transverse coordinates $\mathbf{x}_\perp=\chi\pmb{\beta}$ using the flat sky approximation. Here $\pmb{\beta}$ refers to the angular coordinate on the sky of a physical point $\mathbf{x}$, as seen from the observer. We estimate the dark matter gravitational potential running $N$--body simulations (with $N=512^3$) with the public code Gadget2 \citep{Gadget2}, using a comoving box size of $240h^{-1}$Mpc. Using a similar procedure as in \citep{RayTracingJain,RayTracingHartlap}, the equation that governs the light ray deflections can be written in the form

@@

where $\mathbf{x}(\chi)$ is the trajectory of a single light ray. Suppose that a light ray reaches the observer at an angular position $\pmb{\theta}$ on the sky: we want to know where this light ray originated, knowing it comes from a redshift $z_s$. To answer this question we need to integrate equation (\ref{raytrajectory}) with the initial conditions $\pmb{\beta}(0;\pmb{\theta})=\pmb{\theta}$, $\dot{\pmb{\beta}}(0;\pmb{\theta})=0$ up to a distance $w_s=w(z_s)$ to obtain the source angular position $\pmb{\beta}(w_s;\pmb{\theta})$; for the light ray trajectory solver, based on equation (\ref{raytrajectory}), we use our proprietary implementation Inspector Gadget. Once we know the details of the light ray trajectories, we can easily infer the interesting weak lensing  quantities by taking the angular derivatives of the ray deflections $A(\chi_s;\pmb{\theta}) = \partial \pmb{\beta}(\chi_s;\pmb{\theta})/\partial\pmb{\theta}$ and performing the usual spin decomposition to infer the convergence $\kappa$ and the shear components $(\gamma^1,\gamma^2)$

@@

\section{Statistical methods}
The goal of this section is to describe the framework in which we combine the CFHT data and our simulations in order to derive the constraints on the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$; we measure a set of statistical descriptors from the data and the simulations, which will then be compared in a Bayesian framework in order to compute parameter confidence intervals.

@@

We studied the accuracy of emulator, built with the CFHTemu1 simulations, by interpolating the convergence descriptors to the fiducial parameter setting $(\Omega_m,w,\sigma_8)=(0.26,-1.0,0.8)$ and comparing the result to the one expected from the CFHTcov simulations. Figure \ref{emulatorAccuracy} shows that our Power Spectrum emulator has accuracy better than 10\% for the lower multipoles and comparable to 1\% for the higher multipoles. The Minkowski Functionals emulator has accuracy $\lesssim$10\% for the first 30 bins (which correspond to $\kappa$ values in $[-0.04,0.08]$) and deteriorates due to numerical noise for the remaining 20 bins; we take care of these inaccuracies in our dimensionality reduction framework, which we explain in the next section.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/emulator_accuracy.eps}
\end{center}
\caption{Accuracy of the emulator based on the CFHTemu1 simulations: we show the difference between the interpolated descriptor at the fiducial parameter setting and the expected descriptor from the CFHTcov simulations, in units of the standard deviation in each bin $i$ (determined from the diagonal elements of the CFHTcov covariance matrix). We show the accuracy results for the Power Spectrum (red) and Minkowski Functionals $V_0$ (green), $V_1$ (blue) and $V_2$ (black).}
\label{emulatorAccuracy}
\end{figure}

\subsection{Dimensionality reduction}
\label{pcasection}
The main goal of this work is constraining the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$ using the CFHTLenS data. When the question on the accuracy of the $N\sigma$ contours is considered, particular attention must be paid to the effect of binning choices on contour sizes computed with equations (\ref{parameterlikelihood})--(\ref{ennesigma}). It is known that the choice of the number of bins, $N_b$, can have a non-negligible effect on the contour sizes (see \citep{Petri2013} for an example with simulated datasets); in order for our results to be robust under this effect, we adopt a Principal Component Analysis (PCA) approach. Our main physical motivation for this approach is that, even though we need to specify $N_b$ numbers in order to fully characterize a binned descriptor, we suspect that the majority of the constraining information (of a particular descriptor) is contained in a limited number of linear combinations of its binning. In the framework adopted by \citep{coyote2}, for example, the authors find that the majority of the cosmological information in the matter Power Spectrum is contained in only 5 selected linear combinations of the multipoles. Because of this, we believe that dimensionality reduction techniques such as PCA can help in delivering accurate cosmological constraints using only a limited number of descriptor degrees of freedom.

@@

Looking at the dimensionality issue from a more geometric perspective, we expect our descriptor space to be a 3-dimensional manifold embedded in a $N_b$--dimensional space. The dimensionality reduction problem is equivalent to the accurate reconstruction of the coordinate chart of this descriptor manifold. As outlined in \citep{astroMLText}, the coordinate chart constructed with the PCA projection in (\ref{pcaprojection}) is accurate for reasonably flat descriptor manifolds. When curvature effects become important, more advanced projection techniques (such as Locally Linear Embedding) have to be employed. We have flexibility on the flatness assumption because we allow the number of principal components to be greater than 3.

@@

%
\begin{figure*}
\includegraphics[scale=0.4]{Figures/pca_components.eps}
\caption{Principal Components of the Power Spectrum(red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and the Moments (orange) descriptor spaces; the left plot shows the magnitudes of the PCA eigenvalues $S_i^2$, the right plot shows their cumulative sum. A dashed black line has been drawn corresponding to $n=3$ components}
\label{pcafig}
\end{figure*}
%
%AP: 3 is the dimensionality of our parameter space which may, or may not tell us something about the number of principal components
%
\begin{figure*}
\includegraphics[scale=0.4]{Figures/robustness_pca_Omega_m-sigma8.eps}
\caption{PCA projection dependence of the $1\sigma$ contours in the $(\Omega_m,\sigma_8)$ plane obtained from a mock observation constructed with the CFHTcov simulations; the different panels refer to the descriptors (from left to right, top to bottom) $V_0$, $\partial V_0$(PDF), $V_1$, $V_2$, Power Spectrum and Moments}
\label{robustnessfig}
\end{figure*}
%

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursOmega_m-sigma8_single.eps}
\includegraphics[scale=0.45]{Figures/contoursmockOmega_m-sigma8_single.eps}
\end{center}
\caption{$1\sigma$ constraints on the $(\Omega_m,\sigma_8)$ parameter doublet using the Power Spectrum (red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and Moments (orange) statistics; we show the constraints from the data (left panel) and from a mock observation constructed using the mean of 1000 realizations in CFHTcov simulations as data (right panel). The contours are calculated from the parameter likelihood function $\mathcal{L}$ marginalized over $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{contours3single}
\end{figure*}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursw-Sigma8Om055_single.eps}
\end{center}
\caption{$1\sigma$ constraints on the $(w,\Sigma_8)$ parameter doublet from the CFHTLenS data, obtained with the Power Spectrum (red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and Moments (orange) statistics; we show the constraints from the data. The contours are calculated from the parameter likelihood function $\mathcal{L}$ marginalized over $\Omega_m$. The parentheses near the descriptor label refer to the number of principal components.}
\label{contours3singleRep}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursSigma8Om055_single.eps}
\end{center}
\caption{$\Sigma_8$ parameter likelihood calculated from the CFHTLenS data using the Power Spectrum (red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and Moments (orange) statistics; the likelihood has been marginalized over $\Omega_m$ and $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{likelihoodSi8single}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursSigma8Om055_cross.eps}
\end{center}
\caption{$\Sigma_8$ parameter likelihood calculated from the CFHTLenS data using the PS (red), PS$\times$Moments (green), MFs (blue), MFs$\times$Power Spectrum (black) and MFs$\times$Power Spectrum$\times$Moments (orange); the likelihood function has been marginalized over $\Omega_m$ and $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{likelihoodSi8cross}
\end{figure}

@@

Regarding the parameter biases, our results are in accordance with  \citep{PetriSpurious}, namely, that unaccounted systematics result in larger parameter biases when the constraints are derived from the MFs, and that the LM statistic is less biased. However, for the CFHTLenS data the MFs can effectively constrain the non-degenerate direction in parameter space, $\Sigma_8$, where the amplifying effects of degeneracy are mitigated (Figure \ref{likelihoodSi8single}).

>>>>

\begin{abstract}
Weak gravitational lensing is a potentially powerful probe for cosmology, and in particular for the dark energy (DE) equation of state $w$. Since typical weak lensing fields are non--Gaussian in nature, quadratic statistics, such as angular power spectra, can miss significant cosmological information contained in survey data. In this work we examine constraints on the parameter triplet $(\Omega_m,w,\sigma_8)$ derived from weak lensing Minkowski functionals and moments, using publicly available data from the 154\,deg$^2$ CFHTLenS survey. We utilize a new suite of ray--tracing N-body simulations spanning 91 points in the $(\Omega_m,w,\sigma_8)$ parameter space, replicating the galaxy sky positions, redshift and shape noise in the CFHTLenS catalogs. We then build an emulator that interpolates the simulated descriptor spaces, and use it to compute the parameter likelihood, from which we derive the data constraints. We find that the dimensionality reduction technique Principal Component Analysis helps in stabilizing the constraints with respect to the number of bins used to construct the convergence statistics. Using our full set of statistical descriptors, we constrain $\Sigma_8=\sigma_8(\Omega_m/0.27)^{0.55}$ to a value of $0.75\pm0.04$ ($1\sigma$), well in agreement with the values already quoted in the literature. We find that, while the Minkowski functionals constraints on the $(\Omega_m,\sigma_8)$ doublet suffer from a bias, the moments provide a tight unbiased bound on the parameters, with a substantial contribution coming from quartic moments of gradients.

@@

Weak gravitational lensing is emerging as a promising technique to constrain cosmology. Techniques have been developed to construct cosmic shear fields with shape measurements in large galaxy catalogues. Although the shear two--point function (2PCF) serves as a main cosmological probe (see \citep{CFHTKilbinger} for example), alternative statistics have been shown to increase the amount of cosmological information one can extract from weak lensing fields. Among these, higher moments (\citep{moments1,moments2,moments3,moments4,moments5,moments6}), three--point functions (\citep{3pcf1,3pcf2}), bispectra (\citep{bispectrum1,bispectrum2,bispectrum3,bispectrum4}), peak counts (\citep{peaks1,peaks2,peaks3,peaks4,peaks5,peaks6}) and Minkowski Functionals (\citep{MinkJan,Petri2013}) have shown to improve cosmological constraints in weak lensing analyses.  
In this work we use the publicly available CFHTLenS data, consisting of a catalog of $\approx$4.2 million galaxies, combined with a suite of ray-tracing simulations in 91 different cosmological models to constrain the cosmological parameters $\Omega_m$,$\sigma_8$ and the DE equation of state $w$. The paper is organized as follows: we first give an overview of the CFHTLenS catalogs, summarizing the data reduction techniques adopted. Next, we give a description of our simulation pipeline, including the procedure used to sample the parameter space and outline the ray--tracing algorithm. We call the weak lensing statistical observables - the Power Spectrum, Minkowski Functionals and Moments - "descriptors" throughout the paper. We discuss the calculation of the statistical descriptors, the dimensional reduction method, and the statistical inference framework used. We then outline our results, and the cosmological parameter constraints obtained. To conclude, we discuss our findings and comment on possible future extensions of this analysis.

@@

\subsection{CFHTLenS data reduction}
\label{cfhtdatareduction}
%
The CFHTLenS survey covers four sky patches of 64, 23, 44 and 23 deg$^2$ area, for a total of 154 deg$^2$. The publicly released data consist of a galaxy catalog created using SExtractor \citep{SExtractor}, a photometric redshift estimated with a Bayesian photometric redshift code \citep{PhotoCode} and galaxy shape measurements made using \textit{lensfit} \citep{cfht1,cfht2}. The cosmological parameter inferences have been obtained in \citep{CFHTKilbinger} using the 2PCF. Additionally a number of authors investigated the CFHTLenS constraining power using statistics that go beyond the usual quadratic ones: \citep{CFHTFu} investigated the effect of using three--point correlations as an additional probe for cosmology, while \citep{CFHTMasato} investigated CFHT cosmological constraints and systematic errors using the Minkowski Functionals. The cosmological constraints obtained in this work are based on an irregular--grid sampling of the parameter space, which is an improvement on the Fisher matrix formalism adopted by \citep{CFHTMasato}.

We apply the following cuts to the galaxy catalog: mask$<1$ (see Table B2 in \citep{SExtractor}), redshift $0.2 < z < 1.3$ (see \citep{cfht1}), fitclass = 0 (which requires the object to be a galaxy) and weight $\mathrm{w}>0$ (with larger $\mathrm{w}$ indicating smaller shear measurement uncertainty). Applying these cuts leaves us 4.2$\times10^6$ galaxies, 124.7 deg$^2$ sky coverage, and average galaxy density $n_{gal} \approx 9.3\,\mathrm{arcmin}^{-2}$. The catalog is further reduced by $\sim25\%$ when one rejects fields with non--negligible star--galaxy correlations. These spurious correlations are likely due to imperfect PSF removal, and do not contain cosmological signal. These cuts are consistent with the ones adopted by the CFHTLenS collaboration (see \citep{CFHTFu}). The CFHTLenS galaxy catalog provides us with the sky position $\pmb{\theta}$, the redshift $z(\pmb{\theta})$ and ellipticity $\mathbf{e}(\pmb{\theta})$ of each galaxy, as well as the individual weight factors $w(\pmb{\theta})$ and additive and multiplicative ellipticity corrections $c(\pmb{\theta}), m(\pmb{\theta})$. Because the CFHTLenS fields are irregularly shaped, we first divide them into 13 squares (subfields) to match the shape and $\approx12$ deg$^2$ size of our simulated maps. These square subfield maps are pixelized according to a Gaussian gridding procedure

@@

\subsection{Simulation design}
In this section we give a description of the method we used to sample the parameter space in our simulation effort. We wish to investigate the non--linear dependence of cosmological probes (in this work Minkowski Functionals and Moments of the $\kappa$ field) on the parameter triplet $\mathbf{p}=(\Omega_m,w,\sigma_8)$, while keeping the other relevant parameters $(h,\Omega_b,n_s)$ fixed to the values (0.7, 0.046, 0.96). We sampled the $D$--dimensional ($D=3$ in this case) parameter space using an irregularly spaced grid, designed with a method similar to \citep{coyote2}. The irregular grid design is more efficient, given the available computing resources: a regular parameter grid with the same average spacing between models would require a prohibitively large number of samples. We limit the parameter sampling in a box of corners $\Omega_m\in[0.07,1],\,w\in[-3.0,0],\,\sigma_8\in[0.1,1.5]$ and we map this sampling box $\Pi$ into an hypercube of unit side. We want to contruct an irregularly spaced grid consisting of $N$ points $\mathbf{x}_i\in[0,1]^D$. Let a \textit{design} $\mathcal{D}$ be the set of this irregularly spaced $N$ points: we wish to find an optimal design, in which the points are spread as uniformly as possible inside the box. Following \citep{coyote2}, we choose our optimal design as the minimum of the cost function

@@

We found that for $N=91$ grid points, order of $10^5$ iterations are sufficient to reach an accuracy of $\epsilon\sim10^{-4}$. Once the optimal design $\mathcal{D}_m$ has been found, we can invert the mapping $\Pi\rightarrow[0,1]^3$ to find our simulation parameter sampling $\mathbf{p}_s$, which we show in Table \ref{designtable} and Figure \ref{designfig}.

@@

\caption{$(\Omega_m,w)$ and $(\Omega_m,\sigma_8)$ projections of our the simulation design. The blue points correspond to the CFHTemu1 simulation set, which consists of one $N$--body simulation per point, while the red point corresponds to the CFHTcov simulation set, which is based on 50 independent $N$--body simulations}

@@

For each parameter point on the grid $\mathbf{p}_s$ we run one $N$--body simulation and perform ray tracing through it, as described in \S~\ref{raysim}, to simulate CFHTLenS shear catalogs: this set of simulations will be called CFHTemu1 throughout the rest of this work. Additionally, we run 50 independent $N$--body simulations with a \textit{fiducial} parameter choice $\mathbf{p}_0=(0.26,-1.0,0.8)$ for the purpose of measuring accurately the covariance matrices which will be used for the parameter inferences in \S-\ref{cosmostats}: this set of simulations will be called CFHTcov throughout the rest of this work.

@@

The goal of this section is to give an outline of our simulation pipeline. 
The fluctuations in the matter density field between a source at redshift $z$ and an observer located on Earth will cause small deflections to the trajectories of light rays traveling from the source to the observer. The fluctuations in the matter field are described by the gravitational potential $\Phi(\mathbf{x},z)=\Phi(\mathbf{x}_\perp,\chi(z))$, where we can replace the physical coordinates $\mathrm{x}$ with the comoving distance from the observer $\chi$ and two transverse coordinates $\mathbf{x}_\perp=\chi\pmb{\beta}$ using the flat sky approximation. Here $\pmb{\beta}$ refers to the angular coordinate on the sky of a physical point $\mathbf{x}$, as seen from the observer. We estimate the dark matter gravitational potential running $N$--body simulations (with $N=512^3$) with the public code Gadget2 \citep{Gadget2}, using a comoving box size of $240h^{-1}$Mpc. Using a similar procedure as in \citep{RayTracingJain,RayTracingHartlap}, the equation that governs the light ray deflections can be written in the form

@@

where $\mathbf{x}(\chi)$ is the trajectory of a single light ray. Suppose that a light ray reaches the observer at an angular position $\pmb{\theta}$ on the sky: we want to know where this light ray originated, knowing it comes from a redshift $z_s$. To answer this question we need to integrate equation (\ref{raytrajectory}) with the initial conditions $\pmb{\beta}(0;\pmb{\theta})=\pmb{\theta}$, $\dot{\pmb{\beta}}(0;\pmb{\theta})=0$ up to a distance $w_s=w(z_s)$ to obtain the source angular position $\pmb{\beta}(w_s;\pmb{\theta})$. For the light ray trajectory solver, based on equation (\ref{raytrajectory}), we use our proprietary implementation Inspector Gadget. Once we know the details of the light ray trajectories, we can easily infer the interesting weak lensing  quantities by taking the angular derivatives of the ray deflections $A(\chi_s;\pmb{\theta}) = \partial \pmb{\beta}(\chi_s;\pmb{\theta})/\partial\pmb{\theta}$ and performing the usual spin decomposition to infer the convergence $\kappa$ and the shear components $(\gamma^1,\gamma^2)$

@@

\section{Statistical methods}
The goal of this section is to describe the framework in which we combine the CFHT data and our simulations in order to derive the constraints on the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$. We measure a set of statistical descriptors from the data and the simulations, which will then be compared in a Bayesian framework in order to compute parameter confidence intervals.

@@

We studied the accuracy of emulator, built with the CFHTemu1 simulations, by interpolating the convergence descriptors to the fiducial parameter setting $(\Omega_m,w,\sigma_8)=(0.26,-1.0,0.8)$ and comparing the result to the one expected from the CFHTcov simulations. Figure \ref{emulatorAccuracy} shows that our Power Spectrum emulator has accuracy better than 10\% for the lower multipoles and comparable to 1\% for the higher multipoles. The Minkowski Functionals emulator has accuracy $\lesssim$10\% for the first 30 bins (which correspond to $\kappa$ values in $[-0.04,0.08]$) and deteriorates due to numerical noise for the remaining 20 bins. We take care of these inaccuracies in our dimensionality reduction framework, which we explain in the next section. Nevertheless, we do not expect these inaccuracies to affect our conclusions as Figure \ref{emulatorAccuracy} clearly show that our emulator is able to distinguish a non--fiducial model from the fiducial one within numerical errors.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/emulator_accuracy.eps}
\end{center}
\caption{Accuracy of the emulator based on the CFHTemu1 simulations: we show the difference between the interpolated descriptor at the fiducial parameter setting and the expected descriptor from the CFHTcov simulations, in units of the standard deviation in each bin $i$ (determined from the diagonal elements of the CFHTcov covariance matrix). We show the accuracy results for the Power Spectrum (red) and Minkowski Functionals $V_0$ (green), $V_1$ (blue) and $V_2$ (black). We also show, using dashed lines, the difference between the expected CFHTcov descriptors and the interpolated descriptor at the non--fiducial point $\mathbf{p}=(0.8,-1.0,0.5)$}
\label{emulatorAccuracy}
\end{figure}

\subsection{Dimensionality reduction}
\label{pcasection}
The main goal of this work is constraining the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$ using the CFHTLenS data. When the question on the accuracy of the $N\sigma$ contours is considered, particular attention must be paid to the effect of binning choices on contour sizes computed with equations (\ref{parameterlikelihood})--(\ref{ennesigma}). It is known that the choice of the number of bins, $N_b$, can have a non-negligible effect on the contour sizes (see \citep{Petri2013} for an example with simulated datasets). In order for our results to be robust under this effect, we adopt a Principal Component Analysis (PCA) approach. Our main physical motivation for this approach is that, even though we need to specify $N_b$ numbers in order to fully characterize a binned descriptor, we suspect that the majority of the constraining information (of a particular descriptor) is contained in a limited number of linear combinations of its binning. In the framework adopted by \citep{coyote2}, for example, the authors find that the majority of the cosmological information in the matter Power Spectrum is contained in only 5 selected linear combinations of the multipoles. Because of this, we believe that dimensionality reduction techniques such as PCA can help in delivering accurate cosmological constraints using only a limited number of descriptor degrees of freedom.

@@

Looking at PCA from a geometrical perspective, the dimensionality reduction problem is equivalent to the accurate reconstruction of the coordinate chart of the descriptor manifold. As outlined in \citep{astroMLText}, the coordinate chart constructed with the PCA projection in (\ref{pcaprojection}) is accurate for reasonably flat descriptor manifolds. When curvature effects become important, more advanced projection techniques (such as Locally Linear Embedding) have to be employed.

@@

%
\begin{figure*}
\includegraphics[scale=0.4]{Figures/pca_components.eps}
\caption{Principal Components of the Power Spectrum(red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and the Moments (orange) descriptor spaces. The left plot shows the magnitudes of the PCA eigenvalues $S_i^2$, the right plot shows their cumulative sum. A dashed black line has been drawn corresponding to $n=3$ components}
\label{pcafig}
\end{figure*}
%
%AP: 3 is the dimensionality of our parameter space which may, or may not tell us something about the number of principal components
%
\begin{figure*}
\includegraphics[scale=0.4]{Figures/robustness_pca_Omega_m-sigma8.eps}
\caption{PCA projection dependence of the $1\sigma$ contours in the $(\Omega_m,\sigma_8)$ plane obtained from a mock observation constructed with the CFHTcov simulations. The different panels refer to the descriptors (from left to right, top to bottom) $V_0$, $\partial V_0$(PDF), $V_1$, $V_2$, Power Spectrum and Moments}
\label{robustnessfig}
\end{figure*}
%

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursOmega_m-sigma8_single.eps}
\includegraphics[scale=0.45]{Figures/contoursmockOmega_m-sigma8_single.eps}
\end{center}
\caption{$1\sigma$ constraints on the $(\Omega_m,\sigma_8)$ parameter doublet using the Power Spectrum (red), $V_0$ (green), $V_1$ (blue), $V_2$ (black) and Moments (orange) statistics. We show the constraints from the data (left panel) and from a mock observation constructed using the mean of 1000 realizations in CFHTcov simulations as data (right panel). The contours are calculated from the parameter likelihood function $\mathcal{L}$ marginalized over $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{contours3single}
\end{figure*}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursw-Sigma8Om055_single.eps}
\end{center}
\caption{$1\sigma$ constraints on the $(w,\Sigma_8)$ parameter doublet from the CFHTLenS data, obtained with the Power Spectrum (red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and Moments (orange) statistics. We show the constraints from the data. The contours are calculated from the parameter likelihood function $\mathcal{L}$ marginalized over $\Omega_m$. The parentheses near the descriptor label refer to the number of principal components.}
\label{contours3singleRep}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursSigma8Om055_single.eps}
\end{center}
\caption{$\Sigma_8$ parameter likelihood calculated from the CFHTLenS data using the Power Spectrum (red), $V_0$ (blue), $V_1$ (green), $V_2$ (black) and Moments (orange) statistics. The likelihood has been marginalized over $\Omega_m$ and $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{likelihoodSi8single}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Figures/contoursSigma8Om055_cross.eps}
\end{center}
\caption{$\Sigma_8$ parameter likelihood calculated from the CFHTLenS data using the PS (red), PS$\times$Moments (green), MFs (blue), MFs$\times$Power Spectrum (black) and MFs$\times$Power Spectrum$\times$Moments (orange). The likelihood function has been marginalized over $\Omega_m$ and $w$. The parentheses near the descriptor label refer to the number of principal components.}
\label{likelihoodSi8cross}
\end{figure}

@@

Regarding the parameter biases, our results are in accordance with  \citep{PetriSpurious}, namely, that unaccounted systematics result in larger parameter biases when the constraints are derived from the MFs, and that the LM statistic is less biased. However, for the CFHTLenS data the MFs can effectively constrain the non-degenerate direction in parameter space, $\Sigma_8$ (Figure \ref{likelihoodSi8single}).
